Facetting
====================
- One of the most useful techniques in data visualization is rendering groups of data alongside each other, making it easy to compare the groups. 
- With ggplot2, one way to do this is by mapping a discrete variable to an aesthetic, like x position, color, or shape.
- Another way of doing this is to create a subplot for each group and draw the subplots
side by side.
- **ggplot2** has two ways to do this: `facet_grid()` and `facet_wrap()`.

<hr>
### Facet Grid 
- With `facet_grid()`, you can specify a variable to split the data into vertical subpanels,
and another variable to split it into horizontal subpanels


```{r}
library(ggplot2)
# The base plot
p <- ggplot(mpg, aes(x=displ, y=hwy)) + geom_point()

# Faceted by drv, in vertically arranged subpanels
p + facet_grid(drv ~ .)

# Faceted by cyl, in horizontally arranged subpanels
p + facet_grid(. ~ cyl)

# Split by drv (vertical) and cyl (horizontal)
p + facet_grid(drv ~ cyl)

```
<hr>
### Facet Wraps
- `facet_wrap()` creates and labels a plot for every level of a factor which is passed to it. 
- Its primary argument takes the form of a one sided formula: `~Factor`. 
- It will then try to efficiently "wrap" these plots into a 2d grid.
```{r}
p + facet_wrap( ~ class)
```


```{r}
# These will have the same result: 2 rows and 4 cols
p + facet_wrap( ~ class, nrow=2)
p + facet_wrap( ~ class, ncol=5)
```

<hr>
### Facet Scales

- Usually you will want all of your facets to have the same x and y scales. If you're plotting the same data in each facet, having free scales on each of the facets will ruin comparability across facets. 
- However, sometimes it will be appropriate to have free scales. You can do this by passing `scales = "free"` to `facet_wrap()`.

```{r}
p + facet_wrap(~class, scales = "free")
```  


Facetting with ggplot
===========================================
Facetting
- Split into three subplots for each level of cylinder
- Each row corresponds to a level of cylinder.
```{r}
qplot(mpg, wt, data=mtcars, facets= cyl~.,
geom=c("point", "smooth"))
```
-  Three categories of cylinder : 4, 6 and 8

#### Facetting (example 2)
Using a linear regression model time
```{r}
qplot(mpg, wt, data=mtcars,
 facets=cyl~.,
 geom=c("point", "smooth"), method="lm")
```
#### Diamonds data set
```{r}
head(diamonds)
```

#### Different display of bar plots
```{r}
qplot(clarity, data=diamonds, geom="bar", fill=cut,
position="stack")
``` 
```{r} 
qplot(clarity, data=diamonds, geom="bar", fill=cut,
position="dodge")
``` 
```{r} 
qplot(clarity, data=diamonds, geom="bar", fill=cut,
position="fill")
``` 
```{r} 
qplot(clarity, data=diamonds, geom="bar", fill=cut,
position="identity")
```

1. Stack
2. Dodge
3. Fill
4. Identity 

#### Frequency Polygons
```{r}
qplot(clarity, data=diamonds, geom="freqpoly",
group=cut, colour=cut, position="identity")
``` 
```{r} 
qplot(clarity, data=diamonds, geom="freqpoly",
group=cut, colour=cut, position="stack")
```
#### Contingency table

```{r} 
table(diamonds$cut,diamonds$clarity)

``` 



More on faceting
```{r}
qplot(cty,data=Mpg2,geom="histogram",binwidth=2)
+ facet_grid(cyl ~ drv)
```
#### Margins
Faceting a plot is like creating a contingency table. It is often useful to display marginal totals as well the individual cells

#### Facet Wrap
An alternative to the grid is a wrapped ribbon of plots, `facet_wrap()` generates a long ribbon of plots, and wraps it into 2d.

#### Controlling scales
For both types of faceting you can control whether or not the position scales are the same in all panels, or allowed to vary (i.e. are the scales
“fixed” or “free”?) Fixed allows for us to compare subsets on an equal basis. It is the default setting.
Free allows us to examine each panel at the appropriate level of detail.
To specify “free” scaling, we add the phrase “scales = free” to the
`facet_grid()` or `facet_wrap()` commands.


```{r}
qplot(cty,data=Mpg2,geom="histogram",binwidth=2) +   facet_grid(cyl ~ .,scales = "free")

``` 
