ggplot2 `stat_summary()`
==============================================
One of the statistics, `stat_summary()`, is somewhat special, and merits its own discussion. `stat_summary()` takes a few different arguments.

- `fun.y` A function to produce y aestheticss
- `fun.ymax` A function to produce ymax aesthetics
- `fun.ymin` A function to produce ymin aesthetics
- `fun.data` A function to produce a named vector of aesthetics.

You pass a function to each of these arguments, and ggplot2 will use the value returned by that function for the corresponding aesthetic. If you pass a function to fun.data, you can compute many summary statistics and return them as a vector, where each element in the vector is named for the aesthetic it should be used for.

```{r}
ggplot(diamonds, aes(cut, price)) +
  stat_summary(fun.y = median, geom = "bar")
```

```{r}
median.quartile <- function(x){
	out <- quantile(x, probs = c(0.25,0.5,0.75))
	names(out) <- c("ymin","y","ymax")
	return(out) 
}
				
ggplot(diamonds, aes(cut, price)) +
  stat_summary(fun.data = median.quartile, geom = "pointrange")
```  
