Stats
=======================
*http://sharpstatistics.co.uk/r/ggplot2-guide/*
Stats apply statistical transformations that are used to summarise the data, and allows a huge range of possibilities. Stat_smooth is a nice stat to illustrate the principles, which fits a line and a shaded band to indicate some specified level of uncertainty, as shown in the following example which fits a linear regression line.

```{r}
ggplot(iris,aes(Sepal.Length,Sepal.Width))+geom_point()+ stat_smooth(method="lm")
```
Stats are automatically split according to the factors so

```{r}
ggplot(iris,aes(Sepal.Length,Sepal.Width,color=factor(Species)))+geom_point()+ stat_smooth(method="lm")
```
produces a plot with 3 sets of fitted lines on which could be considered a little confusing. As faceting is similar to using factors the stat is also automatically split over the plots as shown below.
```{r}
p <- ggplot(iris, aes(Sepal.Length, Sepal.Width)) + geom_point()
p + facet_grid(. ~ Species) + stat_smooth(method = "lm")
```

There are various options for the model, which again is listed in the notes on the Ggplot2 website. For example if you want a quadratic model, use the standard R notation. Note that the variables are referred to as x and y not by name.

stat_smooth(method="lm",formula=y~I(x^2))

As the stat is just another layer on the plot the original geom can be omitted if required just to show the stat.

ggplot(iris,aes(Sepal.Length,Sepal.Width,color=factor(Species)))+ stat_smooth(method="lm")

A very flexible stat is the stat_summary function which allows definition of custom summary statistics. As an example to produce a bar chart of the mean value use,

ggplot(iris,aes(Species,Sepal.Width))+stat_summary(fun.y=mean,geom="bar")

which works as mean takes a vector and returns as single value that defines the y value for each bar. Stat_summary can also take 2 other values ymin and ymax which define the upper and lower bounds.The following function shows how to specify your own summaries. Define a function that takes a vector and returns the mean plus and minus 1 standard deviation, and names the two results as ymin and ymax

myFunc = function(x) {
result = c(mean(x) - sd(x), mean(x) + sd(x))
names(result) = c("ymin", "ymax")
result
}

then pass this function into stat_summary and the result is used to define the limits for error bars.

p <- ggplot(iris, aes(Species, Sepal.Length))
p + stat_summary(fun.y = mean, geom = "point", color = "red", size = 5) + stat_summary(fun.data = "myFunc",
geom = "errorbar", width = 0.2)
