Stats
=======================
### Displaying Statistics
You'll frequently want to add statistical analyses to your plots, or your plots may just be of statistical summaries anyway. **ggplot2** has a few built in statistics to make plotting easier.

A frequently used statistic is a smoothing line with `stat_smooth()`. There are a number of different smoothing lines you can add, from local regression lines (loess) to linear or logistic regressions. 

Let's start with the mpg data again.
```{r}
library(ggplot2)
p <- ggplot(mpg, aes(displ, hwy))

p + geom_point() + stat_smooth()
```
By default, `stat_smooth()` has added a loess line with the standard error represented by a semi-transparent ribbon. You could also specify the method to use to add a different smoothing line.
```{r}
p + geom_point() + stat_smooth(method = "lm")
```
```{r}
library(MASS) 
p + geom_point() + stat_smooth(method = "rlm")
```
Now, statistics are represented with default geometries. 
For `stat_smooth()`, its default geoms are `geom_ribbon() + geom_smooth()`. You could also (inadvisedly) represent the output of the smoothing function with points and errorbars.
```{r}
# No a good plot
p + stat_smooth(geom = "point")+stat_smooth(geom = "errorbar")
```
<hr>

- *http://sharpstatistics.co.uk/r/ggplot2-guide/*



- Stats apply statistical transformations that are used to summarise the data, and allows a huge range of possibilities. 

- `Stat_smooth` is a useful stat to illustrate the principles, which fits a line and a shaded band to indicate some specified level of uncertainty, as shown in the following example which fits a linear regression line.

```{r}

ggplot(iris,aes(Sepal.Length,Sepal.Width))+geom_point()+ stat_smooth(method="lm")
```
- Stats are automatically split according to the factors so produces a plot with 3 sets of fitted lines on which could be considered a little confusing. 

```{r}
ggplot(iris,aes(Sepal.Length,Sepal.Width,color=factor(Species)))+geom_point()+ stat_smooth(method="lm")
```
- As faceting is similar to using factors the stat is also automatically split over the plots as shown below.

```{r}
p <- ggplot(iris, aes(Sepal.Length, Sepal.Width)) + geom_point()
p + facet_grid(. ~ Species) + stat_smooth(method = "lm")
```

There are various options for the model, which again is listed in the notes on the Ggplot2 website. For example if you want a quadratic model, use the standard R notation. Note that the variables are referred to as x and y not by name.
```{r}
stat_smooth(method="lm",formula=y~I(x^2))
```
- As the stat is just another layer on the plot the original geom can be omitted if required just to show the stat.
```{r}
ggplot(iris,aes(Sepal.Length,Sepal.Width,color=factor(Species)))+ stat_smooth(method="lm")
```
### stat_summary
A very flexible stat is the `stat_summary` function which allows definition of custom summary statistics. As an example to produce a bar chart of the mean value of Sepal width for each species of iris.
```{r}
mean(iris$Sepal.Width)
ggplot(iris,aes(Species,Sepal.Width))+stat_summary(fun.y=mean,geom="bar")
```
which works as mean takes a vector and returns as single value that defines the y value for each bar. 
- `stat_summary` can also take 2 other values `ymin` and `ymax` which define the upper and lower bounds.
- The following function shows how to specify your own summaries. 
- Define a function that takes a vector and returns the mean plus and minus 1 standard deviation, and names the two results as `ymin` and `ymax`.

```{r}
myFunc = function(x) {
result = c(mean(x) - sd(x), mean(x) + sd(x))
names(result) = c("ymin", "ymax")
result
}
```
then pass this function into `stat_summary`
and the result is used to define the limits for ***error bars***.
```{r}
p <- ggplot(iris, aes(Species, Sepal.Length))
p + stat_summary(fun.y = mean, geom = "point", color = "red", size = 5) + stat_summary(fun.data = "myFunc",
geom = "errorbar", width = 0.2)
```


Sure thing, Kevin. Here's your tutorial tidied up, with clearer headings, improved formatting, and concise explanations for readability and flow:

---

# 📊 ggplot2 Statistical Layers: `stat_` Functions

## Overview
In **ggplot2**, `stat_` functions apply statistical transformations that summarize your data for visualization. These are useful for tasks like trend fitting, summary statistics, and distributions.

---

## 1. `stat_smooth()` – Smoothing Lines
This function overlays a fitted curve (e.g., loess or linear) and a confidence interval ribbon.

```r
library(ggplot2)
p <- ggplot(mpg, aes(displ, hwy))
p + geom_point() + stat_smooth()
```

You can change the method of smoothing:

```r
p + geom_point() + stat_smooth(method = "lm")        # Linear model
p + geom_point() + stat_smooth(method = "rlm")       # Robust linear model (from MASS package)
```

**Custom formula** example:

```r
stat_smooth(method = "lm", formula = y ~ I(x^2))     # Quadratic curve
```

Grouped by factor or faceted:

```r
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) + 
  geom_point() + stat_smooth(method = "lm")

ggplot(iris, aes(Sepal.Length, Sepal.Width)) + 
  geom_point() + facet_grid(. ~ Species) + stat_smooth(method = "lm")
```

You can even remove `geom_point()` to show only the smooth line.

---

## 2. `stat_summary()` – Custom Summaries
Summarizes data (e.g. mean, median, min/max) using user-defined or built-in functions.

### Example: Bar of Species Mean Sepal Width

```r
ggplot(iris, aes(Species, Sepal.Width)) + 
  stat_summary(fun.y = mean, geom = "bar")
```

### Example: Point and Error Bars

Define a function to return ymin and ymax:

```r
myFunc <- function(x) {
  c(ymin = mean(x) - sd(x), ymax = mean(x) + sd(x))
}

ggplot(iris, aes(Species, Sepal.Length)) +
  stat_summary(fun.y = mean, geom = "point", color = "red", size = 5) +
  stat_summary(fun.data = myFunc, geom = "errorbar", width = 0.2)
```

### Example: Custom Median + Quartile Range

```r
median.quartile <- function(x){
  out <- quantile(x, probs = c(0.25, 0.5, 0.75))
  names(out) <- c("ymin", "y", "ymax")
  return(out)
}

ggplot(diamonds, aes(cut, price)) +
  stat_summary(fun.data = median.quartile, geom = "pointrange")
```

---

## 3. `stat_ecdf()` – Empirical CDFs

Shows the proportion of data points less than or equal to each x value.

```r
df <- data.frame(
  x = c(rnorm(100, 0, 3), rnorm(100, 0, 10)),
  g = gl(2, 100)
)

ggplot(df, aes(x, colour = g)) + stat_ecdf()
```

Or with `qplot()`:

```r
qplot(rnorm(1000), stat = "ecdf", geom = "step")
```

---
